Master thesis problem statement: Keyword search over heterogeneous sources

### 2012-10-11 doc/thesis/latex/proposed_solutions.tex
applied valentins patch

	\item {\color{red} evaluate different user feedback approaches (e.g. HMM)
	\item multi-wildcard queries (e.g. dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{Matching to structured query will be harder and guessing even worse. Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants still to be flexible at least for datasets}.}

@@

@@

Also consider:
                			\begin{itemize}
                			%\item given an entity received from provider, determining if it is useful to cache for long-term
%	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
                			%\item could cache even \textit{expired} data, but warning user
                			\item could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
                			\item Can we automatically figure which fields are static and which are changing?
                			\item pre-fetching common (sub-)queries: determining manually and/or automatically
                			\end{itemize}

>>>>

	\item {\color{red} TODO: evaluate different user feedback approaches (e.g. HMM) {\color{green} please elaborate how?}
	\item multi-wildcard queries (e.g. dataset=*Zmm*special*RECO*) are common at CMS that will need special attention\footnote{Matching to structured query will be harder and guessing even worse. Also for performance reasons some web-services will introduce limitations on the format of wilcard queries, while DAS wants still to be flexible at least for datasets}.}

@@

+\item {\color{green} Build series of queries which will answer the given user question, e.g. user asks {\it I want to find files which contains runs taken with magnetic field 4T for given dataset at site X}. This should be decomposed somehow (subject to research) into series of quries, e.g. Find datasets at site X; find files for given dataset list; find runs for found files; filter runs with magnetic field 4T}

@@

Also consider:
                			\begin{itemize}
                			%\item given an entity received from provider, determining if it is useful to cache for long-term
%	                			\item \textbf{\color{red}At what level are we caching now? query or individual query results?}
                			%\item could cache even \textit{expired} data, but warning user
                			\item could also have different validity dates for certain fields. if no volatile fields are not explicitly requested, even a very old cache could be used. 
                			\item Can we automatically figure which fields are static and which are changing?
                			 {\color{green} Yes, if we enfore APIs to record modification timestamp}
                			 {\color{red} I also meant some automatic algorithm, however that would be more complex and less reliable}
                			\item pre-fetching common (sub-)queries: determining manually and/or automatically
                			\end{itemize}

### 2012-10-11 doc/thesis/latex/report_sept.tex
updated problem definition, solutions and literature review. review to be finished
applied valentins patch

For an IR system with wide variety of users, it important to provide an easy to use interface with fairly flat learning curve, while at the same time not loosing support of fairly complex queries.
%
	Even a simple structured query language containing entity names over the mediated schema may seem hard to learn%
	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).}%
	, especially in the beginning. Therefore, it shall be useful to guide new users interactively through the process of building of their query.
%   
   Supporting non-structured keyword queries is also worth investigation as quite many users reported they are missing this Google-like search experience.

@@

\begin{itemize}
						\item Low data providers performance (e.g. on DBS some queries requires joining many large tables)
	     				\item At the moment all queries are put into one pool and has to wait until some threads are available. If a couple of heavy queries were submitted before hand, even very light queries would have to wait long. Explore more advanced Query prioritization 
	     					(e.g. we could have some query cost model based on history or predefined scores per API)

@@

					\item Could displaying Partial Results improve performance?
					\begin{itemize}
					     \item given current APIs to do aggregation over data sources, DAS has to fetch ALL records matching the query instead of only the first page.
%                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
                        \item 
                        \begin{small}
                        {\color{red}e.g. displaying intermediary results of aggregation while still calculating with statistical bounds -- however statistical bounds are much meaningful only if items are well snuffled/randomized -- people could stop the query if they see a non-sense}
                        \end{small}

@@

\subsection{Other}
\subsubsection*{Handling distributed search efficiently}
{\color{red} People may be filtering data items where selection criteria could be present on two separate autonomous sources, meaning with the current APIs available the results would have to be fetched separately from each source and ones not contained in both filtered out only afterwards that is suboptimal.}

@@

Depending on needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account)}

\include{proposed_solutions}

>>>>

For an IR system with wide variety of users, it important to provide an easy to use interface with fairly flat learning curve, while at the same time not loosing support of fairly complex queries.
%
	Even a simple structured query language containing entity names over the mediated schema may seem hard to learn%
%	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} (even though there may exist slight differences in their naming on different data-sources).}%
	\footnote{On the other hand, at CERN the names in  the mediated schema are referring to real-world entities that {\color{red}are fairly consistently named} {\color{green} Yes, one of the task for DAS is to provide consistent naming convention for end-users} (even though there may exist slight differences in their naming on different data-sources).}%	
	, especially in the beginning. Therefore, it shall be useful to guide new users interactively through the process of building of their query.
%   
   Supporting non-structured keyword queries is also worth investigation as quite many users reported they are missing this Google-like search experience.

@@

\begin{itemize}
						\item Bad data providers performance (e.g. on DBS some queries requires joining many very large tables)
	     				\item At the moment all queries are put into one pool and has to wait until some threads are available. If a couple of heavy queries were submitted before hand, even very light queries would have to wait long. Explore more advanced Query prioritization 
	     					(e.g. we could have some query cost model based on history or predefined scores per API)

@@

					\item Could displaying Partial Results improve performance?
					\begin{itemize}
					     \item given current APIs to do aggregation over data sources, DAS has to fetch ALL records matching the query instead of only the first page.
%                        \item ??? do we have cases then multiple APIs of the same service could be called for the same data, can we choose just one???
                        \item 
                        \begin{small}
                        {\color{red}e.g. displaying intermediary results of aggregation while still calculating with statistical bounds -- however statistical bounds are much meaningful only if items are well snuffled/randomized -- people could stop the query if they see a non-sense}
                          {\color{green} Partial results are problematic to construct in DAS, since DAS does the aggregation across multiple dataprovider. To do so DAS needs to fetch all results. Although it would be possible to do partial results for data coming from a single data-source.}                        
                        \end{small}

@@

\subsection{Other}
\subsubsection*{Handling distributed search efficiently}
{\color{red} People may be filtering data items where selection criteria could be present on two separate autonomous sources, meaning with the current APIs available the results would have to be fetched separately from each source and ones not contained in both filtered out only afterwards that is suboptimal.}
+{\color{green} Please elaborate more, I don't really follow.}

@@

Depending on needs it could be either of exploratory or API/Query Forms based approach, where fake APIs are defined in terms of SQL queries (or even better in a simplified form that would generate SQL by taking schema into account)}
+{\color{green} Indeed, an dthat was the topic of research for another DAS student, Liang Dong. The idea was to generalize DBS-QL to any DB back-end and provide common package which will do a job. The work is done over here https://github.com/vkuznet/PyQueryBuilder. Feel free to check it out and we can discuss it more.}
+
\include{proposed_solutions}
